---
id: "bob-swing-arc"
created: "2026-02"
attribution: QRI-2026 (revised from QR-A kid-ring-screen)
use: "Ch 11 Dynamics – pendulum state space (no hand-drawing)"
---

```{r include=FALSE, eval=!"devoirs" %in% (.packages())}
library(devoirs)
library(mosaicCalc)
```

::: {#exr-bob-swing-arc}

**Note:** In this exercise, $\theta = 0°$ is the position where the pendulum hangs **straight down**. Positive $\theta$ is one side, negative $\theta$ the other.

::: {.callout-tip collapse=true}
## Hint: What does "straight down" mean?
Center/bottom of the pendulum's swing.
:::

The figure below shows six different instantaneous configurations for a pendulum as it swings. Each panel is labeled (a)–(f). The curved arrow in each panel shows the velocity of the bob; the length of the arrow is proportional to the speed (velocity × 1 second). In state space, the horizontal axis is $\theta$ (angle in degrees); the vertical axis is angular velocity.

```{r}
#| echo: false
#| warning: false
#| label: fig-pendulums-6
#| layout-ncol: 3
#| fig-cap: "Six pendulum configurations (a)–(f)."
#| fig-subcap:
#| - "a"
#| - "b"
#| - "c"
#| - "d"
#| - "e"
#| - "f"
source("_software/pendulum.R")
draw_pendulum1(pi/8, -0.2)
draw_pendulum1(-pi/4, 0.05)
draw_pendulum1(pi/7, -0.3)
draw_pendulum1(-pi/6, 0.5)
draw_pendulum1(pi/10, -0.1)
draw_pendulum1(-pi/10, +1)
```

**1.** Which frame shows the pendulum where it is nearest the *top* of its swing?

::: {.callout-tip collapse=true}
## Hint
Maximum angle from vertical.
:::

```{mcq}
#| label: bsa-topmost
#| inline: true
1. a
2. b [correct]
3. c
4. d
5. e
6. f
```

**2.** The next figure shows the **flow field** (dynamical rule) for the pendulum in state space: each arrow shows the direction the state moves from that point ($\theta$ on the horizontal axis, velocity on the vertical). **0° = straight down.** Numbered points **1–5** are marked; some match one of the pendulum frames (a)–(f), at least one does not. Use the flow and the pendulum panels to locate each state.

```{r}
#| label: fig-pendulum-flow
#| fig-cap: "Pendulum flow field. Arrows show (dθ/dt, dv/dt). Numbered points 1–5: some match frames (a)–(f), at least one does not."
#| echo: false
#| warning: false
# Pendulum: d(theta)/dt = v, dv/dt = -k*theta with 0 = straight down
k <- 0.03
pend <- makeODE(dp ~ v, dv ~ -k*p)
# Points: 1=frame b (top), 2=frame f (max speed), 3=frame a (right, down), 4=distractor (0,0), 5=frame d
pts <- tibble::tibble(
  p = c(-45, -18, 22.5, 0, -30),
  v = c(0.05, 1, -0.2, 0, 0.5),
  label = as.character(1:5)
)
# Sparse grid and shorter arrows so the flow is readable; domain focused on where the points lie
flow_field(pend, domain(p = -48:32, v = -0.35:1.15), npts = 6, scale = 0.45) |>
  gf_hline(yintercept = 0, linetype = 2, alpha = 0.6, color = "gray40") |>
  gf_vline(xintercept = 0, linetype = 2, alpha = 0.6, color = "gray40") |>
  gf_point(v ~ p, data = pts, size = 6, color = "darkred", inherit = FALSE) |>
  gf_text(v ~ p, label = ~ label, data = pts, hjust = -0.7, vjust = -0.4, size = 5, fontface = "bold", color = "darkred", inherit = FALSE) |>
  gf_labs(x = expression(theta ~ "(deg, 0 = straight down)"), y = "angular velocity") |>
  gf_refine(
    scale_x_continuous(breaks = c(-45, -30, 0, 22),
                       minor_breaks = NULL),
    scale_y_continuous(breaks = c(-0.2, 0, 0.5, 1),
                       minor_breaks = NULL)
  ) |>
  gf_theme(theme_minimal(base_size = 12))
```

**2a.** Which numbered point (1–5) corresponds to **frame b**?

::: {.callout-tip collapse=true}
## Hint
Frame b is the pendulum nearest the top of its swing.
:::

```{mcq}
#| label: bsa-match-b
#| inline: true
1. 1 [correct]
2. 2
3. 3
4. 4
5. 5
```

**2b.** Which numbered point corresponds to **frame f**?

::: {.callout-tip collapse=true}
## Hint
Frame f is the one with the largest speed in the pendulum figure.
:::

```{mcq}
#| label: bsa-match-f
#| inline: true
1. 1
2. 2 [correct]
3. 3
4. 4
5. 5
```

**2c.** Which numbered point does **not** correspond to any of the six pendulum frames (a)–(f)?

::: {.callout-tip collapse=true}
## Hint
One of the numbered points is a state that never appears in the six panels (a)–(f).
:::

```{mcq}
#| label: bsa-distractor
#| inline: true
1. 1
2. 2
3. 3
4. 4 [correct]
5. 5
```

::: {.callout-tip collapse=true}
## Why is 4 correct?
Point 4 is at (0, 0): straight down at rest. None of the six pendulum panels show that state.
:::

**2d.** Which numbered point corresponds to **frame a**?

::: {.callout-tip collapse=true}
## Hint
Frame a shows the bob on the right, with velocity pointing downward.
:::

```{mcq}
#| label: bsa-match-a
#| inline: true
1. 1
2. 2
3. 3 [correct]
4. 4
5. 5
```

:::
<!-- end of exr-bob-swing-arc -->
