---
id: "bob-swing-arc"
created: "2026-02"
attribution: QRI-2026 (revised from QR-A kid-ring-screen)
use: "Ch 11 Dynamics – pendulum state space (no hand-drawing)"
---

```{r include=FALSE, eval=!"devoirs" %in% (.packages())}
library(devoirs)
library(mosaicCalc)
```

::: {#exr-bob-swing-arc}

The figure below shows six different instantaneous configurations for a pendulum as it swings. Each panel is labeled (a)–(f). The curved arrow in each panel shows the velocity of the bob; the length of the arrow is proportional to the speed (velocity × 1 second). The horizontal axis in state space is position $\theta$ (angle in degrees); the vertical axis is angular velocity.

```{r}
#| echo: false
#| warning: false
#| label: fig-pendulums-6
#| layout-ncol: 3
#| fig-cap: "Six pendulum configurations (a)–(f)."
#| fig-subcap:
#| - "a"
#| - "b"
#| - "c"
#| - "d"
#| - "e"
#| - "f"
source("_software/pendulum.R")
draw_pendulum1(pi/8, -0.2)
draw_pendulum1(-pi/4, 0.05)
draw_pendulum1(pi/7, -0.3)
draw_pendulum1(-pi/6, 0.5)
draw_pendulum1(pi/10, -0.1)
draw_pendulum1(-pi/10, +1)
```

**1.** Which frame shows the pendulum where it is nearest the *top* of its swing (maximum angle from vertical)?

```{mcq}
#| label: bsa-topmost
#| inline: true
1. a
2. b [correct]
3. c
4. d
5. e
6. f
```

**2.** The next figure shows the *state space* for the pendulum: horizontal axis = $\theta$ (degrees), vertical axis = angular velocity. Three states from the pendulum figures are plotted and labeled **A**, **B**, and **C** (they correspond to three of the frames (a)–(f), in no particular order).

```{r}
#| label: fig-pendulum-state-space
#| fig-cap: "State space: three states A, B, C from the pendulum frames."
#| echo: false
# A = top of swing (b), B = largest speed (f), C = right side, moving down (a)
pts <- tibble::tibble(
  theta_deg = c(-45, -18, 22.5),
  velocity  = c(0.05, 1, -0.2),
  label = c("A", "B", "C")
)
gf_blank(0 ~ 0) |>
  gf_lims(y = c(-1.2, 1.2), x = c(-55, 35)) |>
  gf_labs(y = "velocity", x = expression(theta ~ "(deg)")) |>
  gf_point(velocity ~ theta_deg, data = pts, size = 4, color = "steelblue") |>
  gf_text(velocity ~ theta_deg, label = ~ label, data = pts, hjust = -0.5, vjust = -0.5, size = 4) |>
  gf_hline(yintercept = 0, linetype = 2, alpha = 0.5) |>
  gf_vline(xintercept = 0, linetype = 2, alpha = 0.5) |>
  gf_theme(theme_minimal())
```

**2a.** Which point (A, B, or C) corresponds to the **top of the swing** (pendulum nearest maximum angle)?

```{mcq}
#| label: bsa-match-b
#| inline: true
1. A [correct]
2. B
3. C
```

**2b.** Which point corresponds to the configuration with the **largest speed** in the pendulum figure?

```{mcq}
#| label: bsa-match-f
#| inline: true
1. A
2. B [correct]
3. C
```

**3.** The *rate of change of state* has two components: $d\theta/dt$ (horizontal) and $d(\text{velocity})/dt$ (vertical). At the **top of the swing** (point A), in which direction does this rate-of-change vector mainly point?

```{mcq}
#| label: bsa-roc-top
1. Left (negative $\theta$ direction).
2. Right (positive $\theta$ direction).
3. Down (negative velocity direction).
4. Up (positive velocity direction). [correct]
5. Both components are zero.
```

:::
<!-- end of exr-bob-swing-arc -->
