---
id: "bob-swing-arc"
created: "2026-02"
attribution: QRI-2026 (revised from QR-A kid-ring-screen)
use: "Ch 11 Dynamics – pendulum state space (no hand-drawing)"
---

```{r include=FALSE, eval=!"devoirs" %in% (.packages())}
library(devoirs)
library(mosaicCalc)
```

::: {#exr-bob-swing-arc}

The figure below shows six different instantaneous configurations for a pendulum as it swings. Each panel is labeled (a)–(f). The curved arrow in each panel shows the velocity of the bob; the length of the arrow is proportional to the speed (velocity × 1 second). The horizontal axis in state space is position $\theta$ (angle in degrees); the vertical axis is angular velocity.

```{r}
#| echo: false
#| warning: false
#| label: fig-pendulums-6
#| layout-ncol: 3
#| fig-cap: "Six pendulum configurations (a)–(f)."
#| fig-subcap:
#| - "a"
#| - "b"
#| - "c"
#| - "d"
#| - "e"
#| - "f"
source("_software/pendulum.R")
draw_pendulum1(pi/8, -0.2)
draw_pendulum1(-pi/4, 0.05)
draw_pendulum1(pi/7, -0.3)
draw_pendulum1(-pi/6, 0.5)
draw_pendulum1(pi/10, -0.1)
draw_pendulum1(-pi/10, +1)
```

**1.** Which frame shows the pendulum where it is nearest the *top* of its swing (maximum angle from vertical)?

```{mcq}
#| label: bsa-topmost
#| inline: true
1. a
2. b [correct]
3. c
4. d
5. e
6. f
```

**2.** The next figure shows the *state space* for the pendulum: horizontal axis = $\theta$ (degrees), vertical axis = angular velocity. Six states are plotted and labeled **A** through **F**. These correspond to the six pendulum frames (a)–(f) above, but in a different order. Your task: match each state-space point to the correct pendulum frame. (The fastest velocity in the pendulum figure is taken to be 1 on the vertical scale.)

```{r}
#| label: fig-pendulum-state-space
#| fig-cap: "State space: six states A–F corresponding to the six pendulum frames (a)–(f)."
#| echo: false
# Positions (theta in deg, velocity): a=(22.5,-0.2), b=(-45,0.05), c=(25.7,-0.3), d=(-30,0.5), e=(18,-0.1), f=(-18,1)
pts <- tibble::tibble(
  theta_deg = c(22.5, -45, 25.7, -30, 18, -18),
  velocity  = c(-0.2, 0.05, -0.3, 0.5, -0.1, 1),
  label = LETTERS[1:6]
)
gf_blank(0 ~ 0) |>
  gf_lims(y = c(-1.2, 1.2), x = c(-55, 35)) |>
  gf_labs(y = "velocity", x = expression(theta ~ "(deg)")) |>
  gf_point(velocity ~ theta_deg, data = pts, size = 5, color = "steelblue") |>
  gf_text(velocity ~ theta_deg, label = ~ label, data = pts, hjust = -0.5, vjust = -0.5, size = 5) |>
  gf_hline(yintercept = 0, linetype = 2, alpha = 0.5) |>
  gf_vline(xintercept = 0, linetype = 2, alpha = 0.5) |>
  gf_theme(theme_minimal())
```

**2a.** Which state-space point (A–F) corresponds to **frame b** (pendulum nearest the top of its swing)?

```{mcq}
#| label: bsa-match-b
#| inline: true
1. A
2. B [correct]
3. C
4. D
5. E
6. F
```

**2b.** Which state-space point corresponds to **frame f** (the frame with the largest speed in the figure)?

```{mcq}
#| label: bsa-match-f
#| inline: true
1. A
2. B
3. C
4. D
5. E
6. F [correct]
```

**2c.** Frame (a) shows the bob on the right side with velocity pointing roughly downward (negative direction). Which point is frame (a)?

```{mcq}
#| label: bsa-match-a
#| inline: true
1. A [correct]
2. B
3. C
4. D
5. E
6. F
```

**3.** The *rate of change of state* at any point has two components: $d\theta/dt$ (horizontal) and $d(\text{velocity})/dt$ (vertical). You can read $d\theta/dt$ from the pendulum: it equals the angular velocity. The vertical component (rate of change of velocity) is the acceleration; for a pendulum, the acceleration is a restoring force toward the vertical.

**3a.** At the state corresponding to the **top of the swing** (point B), in which direction does the rate-of-change vector $(d\theta/dt,\, d(\text{velocity})/dt)$ mainly point?

```{mcq}
#| label: bsa-roc-top
1. Left (negative $\theta$ direction).
2. Right (positive $\theta$ direction).
3. Down (negative velocity direction).
4. Up (positive velocity direction). [correct — velocity is small and increasing as the bob starts to swing down.]
5. Nowhere; both components are zero.
```

**3b.** At point **A** (bob on the right, moving slowly toward the left), what is the sign of $d\theta/dt$ and the sign of $d(\text{velocity})/dt$?

```{mcq}
#| label: bsa-roc-A
1. $d\theta/dt > 0$, $d(\text{velocity})/dt > 0$
2. $d\theta/dt > 0$, $d(\text{velocity})/dt < 0$
3. $d\theta/dt < 0$, $d(\text{velocity})/dt > 0$
4. $d\theta/dt < 0$, $d(\text{velocity})/dt < 0$ [correct — velocity is negative; restoring force pulls toward center so acceleration is negative.]
5. One is zero.
```

:::
<!-- end of exr-bob-swing-arc -->
