---
id: "calf-wear-fork"
created: "Fri Jan 16 15:14:24 2026"
attribution: TBA
---

```{r include=FALSE, eval=!"devoirs" %in% (.packages())}
# For stand-alone compilation testing
library(devoirs)
library(mosaicCalc)
```

::: {#exr-calf-wear-fork}
<!-- this is a subset of zebra-dream-blouse from QR-A -->

The `Engine` data frame, which contains measurements of many different internal combustion engines of widely varying sizes. For instance, we can graph engine RPM (revolutions per second) versus engine mass (in pounds), as in @fig-rpm-mass1.

```{r echo=FALSE}
#| label: fig-rpm-mass1
#| fig-cap: "Engine speed (RPM) versus mass for 39 different engines using *linear axes*. Applications for engines span many magnitudes: compare a small lawn-mower engine to the huge engine in a large ship."
gf_point(RPM ~ mass, data = Engines) |>
  gf_theme(theme_minimal())
```

In the @fig-rpm-mass1, most engines have a mass of... zero. At least, that is how it appears. The scale of the horizontal axis is determined only by the two massive 100,000-pound monster engines plotted at the right end of the graph.

Plotting the magnitude of the engine mass spreads things out, as in @fig-rpm-mass-log.

::: {.column-page-inset-right}
```{r echo=FALSE}
#| label: fig-rpm-mass-log
#| fig-cap: "The engine data plotted with magnitude scales"
#| fig-subcap: 
#| - " "
#| - " "
#| layout-ncol: 2
gf_point(RPM ~ mass, data = Engines) |>
  gf_refine(scale_x_log10()) |>
  gf_theme(theme_minimal())

gf_point(RPM ~ mass, data = Engines) |>
  gf_refine(
    scale_x_log10(),
    scale_y_log10()
    ) |> 
  gf_theme(theme_minimal())
```
:::

1. Which of the plots in @fig-rpm-mass-log uses two magnitude axes?

```{mcq}
#| label: zdb-1-8dw
#| inline: true
1. (a)
2. (b) [correct]
3. both a and b
4. neither a nor b
```

2. In @fig-rpm-mass-log, there is an unlabelled vertical grid line half-way between 100 and 10,000 pounds. What is an appropriate label?

```{mcq}
#| label: zdb-3-k4s
#| inline: true
1. 500
2. 550
3. 1000 [correct]
4. 3000
```

:::
 <!-- end of exr-calf-wear-fork -->
