---
id: "hare-read-chart"
created: "2026-03"
attribution: QRI-2026 (from QR-A rabbit-fight-ring; uses QR-A data)
use: "Chapter 12 – rate of change, CO2 concentration"
---

```{r include=FALSE, eval=!"devoirs" %in% (.packages())}
library(devoirs)
library(mosaicCalc)
```

::: {#exr-hare-read-chart}
"*The global rise in CO_2_ has nearly leveled off.*" --- BBC News, Nov. 6, 2025

The meaning of the above sentence seems plain. Or is it?

1. Which of these do you think is closest to the meaning of the BBC quote?

```{mcq}
#| label: hrc-1
1. The atmospheric concentration of CO_2_ shows nearly no increase now from year to year.
2. The atmospheric concentration of CO_2_ is still growing from year to year. [correct]
3. Neither of the above.
```

The BBC news item is based on measurements of CO_2_ at the Mauna Loa observatory in Hawaii from March 1958 through October 2025. The data are available [here](https://gml.noaa.gov/webdata/ccgg/trends/co2/co2_mm_mlo.txt). The graphic below shows the average measurement each year. The data have been annotated with "slope roses," a visual key for measuring a rate of change. (The units are ppm-per-year.)

```{r}
#| echo: false
#| message: false
#| warning: false
CO2 <- readr::read_csv("../QR-A/www/mauna_loa_CO2.csv")
foo <- CO2 |> summarize(concentration = mean(average), .by = year)
foo |> gf_line(concentration ~ year) |>
  LSTbook::add_slope_rose(keepers = "pos", x = 1958, y = 315) |>
  LSTbook::add_slope_rose(keepers = "pos", x = 2000, y = 370) |>
  gf_line(concentration ~ year) |>
  gf_labs(y = "CO2 Concentration (ppm)", x = "Year") |>
  gf_theme(theme_minimal())
```

2. What is the rate of change of CO_2_ concentration with respect to time in 1970? Pick the closest answer.

```{mcq}
#| label: hrc-2
#| inline: true
1. 0.5 ppm/year
2. 1.0 ppm/year [correct]
3. 1.5 ppm/year
4. 2.0 ppm/year
5. 2.5 ppm/year
```

3. If the rate of change in (2) had held from 1970 through all the intervening time up to 2025, what would the current CO_2_ concentration be? Pick the closest answer.

```{mcq}
#| label: hrc-3
#| inline: true
1. 325 ppm
2. 340 ppm
3. 380 ppm [correct hint]
4. 420 ppm
```

4. In year 2000, what was the rate of change in CO_2_ concentration with respect to time?

```{mcq}
#| label: hrc-4
#| inline: true
1. 0.5 ppm/year
2. 1.0 ppm/year
3. 1.5 ppm/year
4. 2.0 ppm/year [correct]
5. 2.5 ppm/year
```

5. Estimate the rate of change in CO_2_ in 2020-2025.

```{mcq}
#| label: hrc-5
#| inline: true
1. 0.5 ppm/year
2. 1.0 ppm/year
3. 1.5 ppm/year
4. 2.0 ppm/year
5. 2.5 ppm/year [correct]
```

6. Explain, as best you can, how the BBC quote is reflected in the actual data.

`r devoirs::devoirs_text("hrc-bbc")`

:::
<!-- end of exr-hare-read-chart -->
