---
id: "heron-catch-fish"
created: "2026-02"
attribution: QRI-2026
use: "Ch 7/9/10 ICBM prior–posterior example"
---

```{r include=FALSE, eval=!"devoirs" %in% (.packages())}
library(devoirs)
library(mosaicCalc)
```

::: {#exr-heron-catch-fish}

Prior and posterior: the ICBM-style example (from class and the book)

This exercise revisits the example from class and the book in which we learn an unknown probability of success \(p\) by starting with a prior distribution and updating with data to get a posterior distribution.

We use a uniform prior: before seeing any data, we treat every value of \(p\) between 0 and 1 as equally plausible. When we observe \(k\) successes in \(n\) trials, Bayesian updating gives a posterior distribution: the density is highest where the data suggest \(p\) is likely to be.

**Part 1: Flat prior, then 4 consecutive successes**

The figure below shows the prior (no data) and then the posterior after 1 success in 1 try, 2 in 2, 3 in 3, and 4 in 4. The red dashed line in each posterior plot shows the fraction of successes so far (\(k/n\)).

```{r}
#| label: fig-hcf-part1
#| fig-cap: "Prior (no data) and posteriors for 1/1, 2/2, 3/3, 4/4 successes"
#| fig-width: 8
#| fig-height: 10
#| echo: false
plot_posterior <- function(k, n, main = NULL) {
  p_seq <- seq(0, 1, length.out = 300)
  alpha_post <- k + 1
  beta_post <- n - k + 1
  dens <- dbeta(p_seq, alpha_post, beta_post)
  if (is.null(main)) main <- if (n == 0) "Prior (no data)" else paste("Posterior: ", k, " successes in ", n, " trials", sep = "")
  plot(p_seq, dens, type = "l", col = "steelblue", lwd = 2, xlab = "Probability of success (p)", ylab = "Density", main = main, bty = "l")
  if (n > 0) abline(v = k/n, col = "red", lty = 2, lwd = 1.5)
}
par(mfrow = c(3, 2), bty = "l")
for (ii in 0:4) plot_posterior(k = ii, n = ii)
par(mfrow = c(1, 1))
```

**Part 2: Informative prior from past data, then 4 new successes**

Sometimes we already have past data. Suppose we had 7 successes in 16 trials and we use that to form our prior (no longer flat). Then we run 4 more trials and get 4 successes (11 successes in 20 trials total). The figure below shows the prior (based on 7/16) and the posterior after updating with the new data (11/20).

```{r}
#| label: fig-hcf-part2
#| fig-cap: "Prior from 7 successes in 16 trials; posterior after 4 more successes (11/20 total)"
#| fig-width: 8
#| fig-height: 4
#| echo: false
p_seq <- seq(0, 1, length.out = 300)
prior_dens <- dbeta(p_seq, 7 + 1, 16 - 7 + 1)
post_dens  <- dbeta(p_seq, 11 + 1, 20 - 11 + 1)
par(mfrow = c(1, 2), bty = "l")
plot(p_seq, prior_dens, type = "l", col = "steelblue", lwd = 2, xlab = "Probability of success (p)", ylab = "Density", main = "Prior (7 successes in 16 trials)", bty = "l")
plot(p_seq, post_dens, type = "l", col = "darkgreen", lwd = 2, xlab = "Probability of success (p)", ylab = "Density", main = "Posterior (11 successes in 20 trials)", bty = "l")
abline(v = 11/20, col = "red", lty = 2, lwd = 1.5)
par(mfrow = c(1, 1))
```

**Part 3: More data — which posterior is correct?**

Suppose we had 4 successes in 4 trials (as in Part 1). We then run 4 more trials and get 3 successes (7 successes in 8 trials total). The figure below shows four possible posteriors labeled A, B, C, and D. Only one of them is the correct posterior after 7 successes in 8 trials.

```{r}
#| label: fig-hcf-choose
#| fig-cap: "Four candidate posteriors (A–D). Which one is correct for 7 successes in 8 trials?"
#| fig-width: 8
#| fig-height: 8
#| echo: false
p_seq <- seq(0, 1, length.out = 300)
par(mfrow = c(2, 2), bty = "l")
# A: 8/8 (wrong - all successes)
dens_a <- dbeta(p_seq, 9, 1)
plot(p_seq, dens_a, type = "l", col = "steelblue", lwd = 2, xlab = "p", ylab = "Density", main = "A", bty = "l")
abline(v = 1, col = "red", lty = 2, lwd = 1.5)
# B: 7/8 (correct)
dens_b <- dbeta(p_seq, 8, 2)
plot(p_seq, dens_b, type = "l", col = "steelblue", lwd = 2, xlab = "p", ylab = "Density", main = "B", bty = "l")
abline(v = 7/8, col = "red", lty = 2, lwd = 1.5)
# C: 4/8 (wrong)
dens_c <- dbeta(p_seq, 5, 5)
plot(p_seq, dens_c, type = "l", col = "steelblue", lwd = 2, xlab = "p", ylab = "Density", main = "C", bty = "l")
abline(v = 1/2, col = "red", lty = 2, lwd = 1.5)
# D: 2/8 (wrong)
dens_d <- dbeta(p_seq, 3, 7)
plot(p_seq, dens_d, type = "l", col = "steelblue", lwd = 2, xlab = "p", ylab = "Density", main = "D", bty = "l")
abline(v = 2/8, col = "red", lty = 2, lwd = 1.5)
par(mfrow = c(1, 1))
```

**Comprehension questions**

1. In Part 1, the first plot is "Prior (no data)." What does that plot represent?

```{mcq}
#| label: hcf-prior
1. The probability of success after we see the data.
2. Our uncertainty about the probability of success *before* we see any data. [correct]
3. The likelihood of the data given each value of p.
4. The number of successes we expect in the next trial.
```

2. As we go from 1/1 to 2/2 to 3/3 to 4/4 in Part 1, what happens to the posterior?

```{mcq}
#| label: hcf-sharpen
1. It stays flat (uniform).
2. It becomes more concentrated (narrower) and shifts toward the observed success rate. [correct]
3. It becomes more spread out.
4. It always peaks at 1.
```

3. In Part 2, we use a prior based on 7 successes in 16 trials, then see 4 more successes in 4 trials (11/20 total). Why is the posterior in Part 2 not as concentrated near 1 as the "4 successes in 4 trials" posterior in Part 1?

```{mcq}
#| label: hcf-informative
1. Because we used a different prior; the earlier 7/16 pulls the posterior away from 1 and reflects that we have mixed evidence overall (11/20 = 0.55). [correct]
2. Because we ran fewer trials in Part 2.
3. Because we have more total trials in Part 2, so the distribution is wider.
4. Because the success rate is different.
```

4. If we had started with a uniform prior and then observed 11 successes in 20 trials, we would get a posterior centered near 11/20. How would that posterior compare to the one in Part 2 (which used a 7/16 prior then 11/20)?

```{mcq}
#| label: hcf-compare
1. They would be identical.
2. The uniform-prior posterior would be more concentrated near 11/20, since we didn't "mix in" the earlier 7/16. [correct]
3. The uniform-prior posterior would be flatter.
4. The Part 2 posterior would always be to the right of the uniform-prior posterior.
```

5. In Part 3, we had 4 successes in 4 trials, then 3 more successes in 4 more trials (7 in 8 total). Which panel (A–D) in the figure shows the correct posterior for 7 successes in 8 trials?

```{mcq}
#| label: hcf-choose-plot
1. A
2. B [correct]
3. C
4. D
```

6. Suppose we had 2 successes in 4 trials. We then run 2 more trials and get 2 more successes (4 successes in 6 trials total). What happens to the posterior?

```{mcq}
#| label: hcf-more-data
1. It stays the same as the posterior after 2 in 4.
2. It becomes more concentrated and shifts toward 4/6 (about 0.67), since we have more data and the success rate is 4/6. [correct]
3. It becomes more spread out.
4. It shifts toward 0.5 and becomes flatter.
```

7. We have a prior based on 7 successes in 16 trials (as in Part 2). We then see 2 more trials and get 2 successes (9 successes in 18 trials total). How does this new posterior (9/18) compare to the posterior in Part 2 (11/20)?

```{mcq}
#| label: hcf-update-compare
1. The new posterior is centered at 0.5 (9/18), since the new data (2/2) brought the overall success rate down from 11/20 toward 50%. [correct]
2. The new posterior is identical to the 11/20 posterior.
3. The new posterior is centered at 0.55, same as 11/20.
4. The new posterior is centered near 1, because we saw 2 successes in 2 trials.
```

:::
<!-- end of exr-heron-catch-fish -->
