---
id: "lynx-eat-hare"
created: "2026-02"
attribution: QRI-2026
use: "Ch 11 Dynamics – predator-prey (Lotka-Volterra)"
---

```{r include=FALSE, eval=!"devoirs" %in% (.packages())}
library(devoirs)
library(mosaicCalc)
```

::: {#exr-lynx-eat-hare}

**Predator–prey (Lotka–Volterra)**  
State $(R, F)$: prey and predator populations. The flow field shows how the state changes; trajectories can **cycle** around an equilibrium. The figure shows the flow and **two** trajectories (green, brown).

```{r}
#| label: fig-lynx-hare-flow
#| fig-cap: "Predator-prey flow field and two trajectories (green, brown). R = prey, F = predators."
#| echo: false
#| warning: false
#| message: false
r <- 0.6
a <- 0.002
b <- 0.0015
m <- 0.6
LV <- makeODE(dR ~ r*R - a*R*F, dF ~ b*R*F - m*F)
# Two trajectories; integrate long enough for full closed orbits
sol1 <- integrateODE(LV, domain(t = 0:50), R = 600, F = 150)
sol2 <- integrateODE(LV, domain(t = 0:50), R = 200, F = 450)
# Widen domain and limits so both full orbits fit (no clipping)
R_range <- c(80, 720)
F_range <- c(80, 580)
flow_field(LV, domain(R = R_range[1]:R_range[2], F = F_range[1]:F_range[2]), npts = 12) |>
  traj_plot(F(t) ~ R(t), sol2, color = "brown", nt = 0, inherit = FALSE, linewidth = 1.3, alpha = 0.85) |>
  traj_plot(F(t) ~ R(t), sol1, color = "forestgreen", nt = 0, inherit = FALSE, linewidth = 1.9, alpha = 0.95) |>
  gf_lims(x = R_range, y = F_range) |>
  gf_labs(x = "R (prey)", y = "F (predators)") |>
  gf_refine(coord_fixed()) |>
  gf_theme(theme_minimal())
```

**Questions**

1. Where there are many prey and few predators (high $R$, low $F$), which way does the flow mainly point?

```{mcq}
#| label: leh-highR-lowF
1. Mainly to the right (R increasing, F not changing much).
2. Mainly upward (F increasing, R not changing much).
3. Roughly northeast (both R and F increasing). [correct]
4. Mainly downward (F decreasing).
5. Roughly southwest (both R and F decreasing).
```

2. Where there are few prey and many predators (low $R$, high $F$), which way does the flow mainly point?

```{mcq}
#| label: leh-lowR-highF
1. Mainly to the right (R increasing).
2. Roughly southwest (both R and F decreasing). [correct]
3. Roughly northeast (both R and F increasing).
4. Mainly upward (F increasing).
5. Mainly to the left (R decreasing only).
```

3. What shape do the two trajectories have?

```{mcq}
#| label: leh-shape
1. They head straight toward the center and stop.
2. They spiral inward toward a single point.
3. They form closed loops (cycles) that repeat. [correct]
4. They head outward without bound.
5. They are straight lines.
```

4. How does long-term behavior here differ from SIR?

```{mcq}
#| label: leh-vs-SIR
1. SIR has no state space; predator-prey does.
2. SIR trends one way; here the state can cycle and not settle. [correct]
3. SIR has two state variables; predator-prey has only one.
4. Predator-prey has no equilibrium; SIR always has one.
```

5. Starting with many prey and few predators, what typically happens to $F$ (predators) next?

```{mcq}
#| label: leh-fox-increase
1. $F$ only decreases, because there are too many prey.
2. $F$ increases (more food for predators), then later can decrease as prey are eaten down. [correct]
3. $F$ stays constant.
4. $F$ decreases to zero.
```

6. Near the center of the flow field the arrows are short. What does that mean?

```{mcq}
#| label: leh-short-arrows
1. Time runs more slowly there.
2. State changes slowly there; it's the **equilibrium**. [correct]
3. Prey and predators are equal in number.
4. The trajectory cannot pass through that region.
```

7. If the system started exactly at that center point, what would happen?

```{mcq}
#| label: leh-center-behavior
1. Both populations would immediately explode.
2. The state would stay there; neither population would grow nor decline because the rate of change is zero at that point. [correct]
3. The state would spiral inward to zero.
4. The state would move in a straight line away from the center.
```

8. When would the cycle break (populations no longer follow a repeating cycle)?

```{mcq}
#| label: leh-cycle-break
1. If we start with too many prey.
2. One population goes to zero (extinction); the closed orbit is lost. [correct]
3. If we start exactly at the equilibrium.
4. If the cycle runs for more than one full loop.
```

9. (Optional) Why does the flow point roughly southwest when there are few prey and many predators?

`r devoirs::devoirs_text("leh-explain", rows = 3)`

:::
<!-- end of exr-lynx-eat-hare -->
