---
id: "lynx-eat-hare"
created: "2026-02"
attribution: QRI-2026
use: "Ch 11 Dynamics – predator-prey"
---

```{r include=FALSE, eval=!"devoirs" %in% (.packages())}
library(devoirs)
library(mosaicCalc)
```

::: {#exr-lynx-eat-hare}

**Predator–prey**  
State $(R, F)$: prey and predator populations. The flow field shows how the state changes; trajectories can **cycle** around an equilibrium. The figure shows the flow and one trajectory (green).

```{r}
#| label: fig-lynx-hare-flow
#| fig-cap: "Predator-prey flow field and one trajectory (green). R = prey, F = predators."
#| echo: false
#| warning: false
#| message: false
library(ggplot2)

# Parameters; equilibrium at R* = 400, F* = 300
a <- 1.5    # prey growth rate
b <- 0.005  # predation rate
c <- 0.5    # conversion efficiency
d <- 1.0    # predator death rate

dRdt <- function(R, F) a*R - b*R*F
dFdt <- function(R, F) c*b*R*F - d*F

# RK4 trajectory
dt     <- 0.005
steps  <- 20000
R_traj <- numeric(steps)
F_traj <- numeric(steps)
R_traj[1] <- 250
F_traj[1] <- 200

for (i in 2:steps) {
  R <- R_traj[i-1]; F <- F_traj[i-1]
  k1r <- dRdt(R, F);                     k1f <- dFdt(R, F)
  k2r <- dRdt(R+dt/2*k1r, F+dt/2*k1f); k2f <- dFdt(R+dt/2*k1r, F+dt/2*k1f)
  k3r <- dRdt(R+dt/2*k2r, F+dt/2*k2f); k3f <- dFdt(R+dt/2*k2r, F+dt/2*k2f)
  k4r <- dRdt(R+dt*k3r,   F+dt*k3f);   k4f <- dFdt(R+dt*k3r,   F+dt*k3f)
  R_traj[i] <- R + dt/6*(k1r + 2*k2r + 2*k3r + k4r)
  F_traj[i] <- F + dt/6*(k1f + 2*k2f + 2*k3f + k4f)
}

traj <- data.frame(R = R_traj, F = F_traj)

# Flow field grid
R_grid <- seq(150, 750, by = 75)
F_grid <- seq(0, 700, by = 75)
grid   <- expand.grid(R = R_grid, F = F_grid)
grid$dR <- mapply(dRdt, grid$R, grid$F)
grid$dF <- mapply(dFdt, grid$R, grid$F)

mag       <- sqrt(grid$dR^2 + grid$dF^2)
mag[mag == 0] <- 1
arrow_len <- 30
grid$dRn  <- grid$dR / mag * arrow_len
grid$dFn  <- grid$dF / mag * arrow_len

ggplot() +
  geom_segment(
    data = grid,
    aes(x = R, y = F, xend = R + dRn, yend = F + dFn),
    arrow = arrow(length = unit(0.15, "cm"), type = "closed"),
    linewidth = 0.35,
    color = "black"
  ) +
  geom_path(
    data = traj,
    aes(x = R, y = F),
    color = "#22aa22",
    linewidth = 1.6
  ) +
  scale_x_continuous(
    name   = "R (prey)",
    breaks = c(200, 400, 600, 800)
  ) +
  scale_y_continuous(
    name   = "F (predators)",
    breaks = c(0, 100, 200, 300, 400, 500, 600, 700)
  ) +
  coord_cartesian(xlim = c(100, 800), ylim = c(0, 700)) +
  theme_bw(base_size = 13) +
  theme(panel.grid.minor = element_blank())
```

**Questions**

1. Where there are many prey and few predators (high $R$, low $F$), which way does the flow mainly point?

```{mcq}
#| label: leh-highR-lowF
1. Mainly to the right (R increasing, F not changing much).
2. Mainly upward (F increasing, R not changing much).
3. Roughly northeast (both R and F increasing). [correct]
4. Mainly downward (F decreasing).
5. Roughly southwest (both R and F decreasing).
```

2. Where there are few prey and many predators (low $R$, high $F$), which way does the flow mainly point?

```{mcq}
#| label: leh-lowR-highF
1. Mainly to the right (R increasing).
2. Roughly southwest (both R and F decreasing). [correct]
3. Roughly northeast (both R and F increasing).
4. Mainly upward (F increasing).
5. Mainly to the left (R decreasing only).
```

3. What shape does the trajectory have?

```{mcq}
#| label: leh-shape
1. It heads straight toward the center and stops.
2. It spirals inward toward a single point.
3. It forms a closed loop (cycle) that repeats. [correct]
4. It heads outward without bound.
5. It is a straight line.
```

4. How does long-term behavior here differ from SIR?

```{mcq}
#| label: leh-vs-SIR
1. SIR has no state space; predator-prey does.
2. SIR trends one way; here the state can cycle and not settle. [correct]
3. SIR has two state variables; predator-prey has only one.
4. Predator-prey has no equilibrium; SIR always has one.
```

5. Starting with many prey and few predators, what typically happens to $F$ (predators) next?

```{mcq}
#| label: leh-fox-increase
1. $F$ only decreases, because there are too many prey.
2. $F$ increases (more food for predators), then later can decrease as prey are eaten down. [correct]
3. $F$ stays constant.
4. $F$ decreases to zero.
```

6. Near the center of the flow field the arrows are short. What does that mean?

```{mcq}
#| label: leh-short-arrows
1. Time runs more slowly there.
2. State changes slowly there; it's the **equilibrium**. [correct]
3. Prey and predators are equal in number.
4. The trajectory cannot pass through that region.
```

7. If the system started exactly at that center point, what would happen?

```{mcq}
#| label: leh-center-behavior
1. Both populations would immediately explode.
2. The state would stay there; neither population would grow nor decline because the rate of change is zero at that point. [correct]
3. The state would spiral inward to zero.
4. The state would move in a straight line away from the center.
```

8. When would the cycle break (populations no longer follow a repeating cycle)?

```{mcq}
#| label: leh-cycle-break
1. If we start with too many prey.
2. One population goes to zero (extinction); the closed orbit is lost. [correct]
3. If we start exactly at the equilibrium.
4. If the cycle runs for more than one full loop.
```

9. Why does the flow point roughly southwest when there are few prey and many predators?

`r devoirs::devoirs_text("leh-explain", rows = 3)`

:::
<!-- end of exr-lynx-eat-hare -->
