---
id: "lynx-eat-hare"
created: "2026-02"
attribution: QRI-2026
use: "Ch 11 Dynamics – predator-prey (Lotka-Volterra)"
---

```{r include=FALSE, eval=!"devoirs" %in% (.packages())}
library(devoirs)
library(mosaicCalc)
```

::: {#exr-lynx-eat-hare}

**A new dynamic system: predator and prey**

In the SIR model, the state (susceptibles, infectives) moves in one general direction over time. Many two-variable dynamical systems behave differently. Here we introduce a simple **predator–prey** model (often called Lotka–Volterra). The state is $(R, F)$: $R$ = prey population and $F$ = predator population (e.g. in units per square km, or hundreds of individuals).

- When there are many prey and few predators, prey tend to increase and predators can grow (more food).
- When there are many predators and few prey, predators decline (not enough food) and prey can recover.
- The dynamical rule has both $R$ and $F$ increasing or decreasing depending on the current state, and trajectories often **cycle** around an equilibrium instead of heading one way.

The figure below shows the **flow field** (dynamical rule) for such a model, with axes in population units. The domain extends toward the axes (low $R$ or low $F$) so you can see how the flow behaves near those edge cases. Each arrow shows the direction and relative rate of change of the state. Two **trajectories** are drawn: one starting at (600, 150) and one at (200, 450). Time runs along each curve; one full cycle might represent several years in a real system.

```{r}
#| label: fig-lynx-hare-flow
#| fig-cap: "Predator-prey flow field and two trajectories. Domain extends toward R=0 and F=0 so edge cases (e.g. near-extinction) are visible. R = prey, F = predators (population units)."
#| echo: false
#| warning: false
#| message: false
# Lotka-Volterra: dR/dt = r*R - a*R*F, dF/dt = b*R*F - m*F
# Equilibrium R* = m/b, F* = r/a. Parameters chosen so equilibrium ~(400, 300), cycle period ~several time units
r <- 0.6
a <- 0.002
b <- 0.0015
m <- 0.6
LV <- makeODE(dR ~ r*R - a*R*F, dF ~ b*R*F - m*F)
sol1 <- integrateODE(LV, domain(t = 0:35), R = 600, F = 150)
sol2 <- integrateODE(LV, domain(t = 0:35), R = 200, F = 450)
# Extend domain toward R=0 and F=0 so edge cases (extinction, cycle break) are visible
flow_field(LV, domain(R = 15:700, F = 15:500), npts = 12) |>
  traj_plot(F(t) ~ R(t), sol1, color = "darkgreen", nt = 0, inherit = FALSE, linewidth = 1.2, alpha = 0.8) |>
  traj_plot(F(t) ~ R(t), sol2, color = "brown", nt = 0, inherit = FALSE, linewidth = 1.2, alpha = 0.8) |>
  gf_labs(x = "R (prey)", y = "F (predators)") |>
  gf_refine(coord_fixed()) |>
  gf_theme(theme_minimal())
```

**Questions**

1. At the point in state space where there are *many* prey and *few* predators (e.g. high $R$, low $F$ — right side of the plot), in which direction does the flow mainly point?

```{mcq}
#| label: leh-highR-lowF
1. Mainly to the right (R increasing, F not changing much).
2. Mainly upward (F increasing, R not changing much).
3. Roughly northeast (both R and F increasing). [correct]
4. Mainly downward (F decreasing).
5. Roughly southwest (both R and F decreasing).
```

2. At the point where there are *few* prey and *many* predators (e.g. low $R$, high $F$ — left side of the plot), in which direction does the flow mainly point?

```{mcq}
#| label: leh-lowR-highF
1. Mainly to the right (R increasing).
2. Roughly southwest (both R and F decreasing). [correct]
3. Roughly northeast (both R and F increasing).
4. Mainly upward (F increasing).
5. Mainly to the left (R decreasing only).
```

3. What shape do the two trajectories (green and brown) in the figure have?

```{mcq}
#| label: leh-shape
1. They head straight toward the center and stop.
2. They spiral inward toward a single point.
3. They form closed loops (cycles) that repeat. [correct]
4. They head outward without bound.
5. They are straight lines.
```

4. How does this predator–prey system *differ* from the SIR epidemic model in terms of long-term behavior?

```{mcq}
#| label: leh-vs-SIR
1. SIR has no state space; predator-prey does.
2. In SIR the state eventually moves in one general direction (e.g. toward fewer infectives); in this predator-prey model the state can cycle and not settle at a single point. [correct]
3. SIR has two state variables; predator-prey has only one.
4. Predator-prey has no equilibrium; SIR always has one.
```

5. Suppose we start at a state with many prey and few predators. Over the next stretch of time, what typically happens to the predator population $F$?

```{mcq}
#| label: leh-fox-increase
1. $F$ only decreases, because there are too many prey.
2. $F$ increases (more food for predators), then later can decrease as prey are eaten down. [correct]
3. $F$ stays constant.
4. $F$ decreases to zero.
```

6. In the **center** of the flow field (e.g. near $R \approx 400$, $F \approx 300$), the arrows are relatively short. What does that indicate?

```{mcq}
#| label: leh-short-arrows
1. Time runs more slowly there.
2. The rate of change of the state is small there (state is changing slowly); this is the **equilibrium** where inflows and outflows balance. [correct]
3. Prey and predators are equal in number.
4. The trajectory cannot pass through that region.
```

7. What would happen if the system started *exactly* at that center point (the equilibrium)?

```{mcq}
#| label: leh-center-behavior
1. Both populations would immediately explode.
2. The state would stay there; neither population would grow nor decline because the rate of change is zero at that point. [correct]
3. The state would spiral inward to zero.
4. The state would move in a straight line away from the center.
```

8. In what kind of deviation or situation would the **cycle break** — that is, the populations would no longer follow a repeating cycle?

```{mcq}
#| label: leh-cycle-break
1. If we start with too many prey.
2. If one population is driven to zero (e.g. extinction of prey or predators); the system then leaves the regime where both species interact and the closed orbit is lost. [correct]
3. If we start exactly at the equilibrium.
4. If the cycle runs for more than one full loop.
```

9. (Optional) In one or two sentences, explain in terms of prey and predators why the flow points roughly southwest when there are few prey and many predators.

`r devoirs::devoirs_text("leh-explain", rows = 3)`

:::
<!-- end of exr-lynx-eat-hare -->
