---
id: "lynx-eat-hare"
created: "2026-02"
attribution: QRI-2026
use: "Ch 11 Dynamics – predator-prey (Lotka-Volterra)"
---

```{r include=FALSE, eval=!"devoirs" %in% (.packages())}
library(devoirs)
library(mosaicCalc)
```

::: {#exr-lynx-eat-hare}

**A new dynamic system: predator and prey**

In the SIR model, the state (susceptibles, infectives) moves in one general direction over time. Many two-variable dynamical systems behave differently. Here we introduce a simple **predator–prey** model (often called Lotka–Volterra). The state is $(R, F)$: $R$ = number of rabbits (prey) and $F$ = number of foxes (predators).

- When there are many rabbits and few foxes, rabbits tend to increase (and foxes can grow by eating them).
- When there are many foxes and few rabbits, foxes tend to decrease (not enough food) and rabbits can recover.
- The dynamical rule therefore has both $R$ and $F$ increasing or decreasing depending on the current state, and trajectories can **cycle** around an equilibrium instead of heading one way.

The figure below shows the **flow field** (dynamical rule) for one such predator–prey model. The horizontal axis is $R$ (rabbits) and the vertical axis is $F$ (foxes). Each arrow shows the direction and relative rate of change of the state at that point. Two **trajectories** are drawn: one starting at (2, 1.5) and one starting at (1, 3).

```{r}
#| label: fig-lynx-hare-flow
#| fig-cap: "Predator-prey flow field and two trajectories. R = rabbits (prey), F = foxes (predators)."
#| echo: false
#| warning: false
#| message: false
# Lotka-Volterra: dR/dt = r*R - a*R*F, dF/dt = b*R*F - m*F
# Equilibrium: R = m/b, F = r/a. Use r=1, a=0.5, b=0.3, m=0.4 => (R*,F*) ~ (1.33, 2)
r <- 1
a <- 0.5
b <- 0.3
m <- 0.4
LV <- makeODE(dR ~ r*R - a*R*F, dF ~ b*R*F - m*F)
sol1 <- integrateODE(LV, domain(t = 0:25), R = 2, F = 1.5)
sol2 <- integrateODE(LV, domain(t = 0:25), R = 1, F = 3)
flow_field(LV, domain(R = 0.4:3, F = 0.4:3.5), npts = 12) |>
  traj_plot(F(t) ~ R(t), sol1, color = "darkgreen", nt = 0, inherit = FALSE, linewidth = 1.2, alpha = 0.8) |>
  traj_plot(F(t) ~ R(t), sol2, color = "brown", nt = 0, inherit = FALSE, linewidth = 1.2, alpha = 0.8) |>
  gf_labs(x = "R (rabbits)", y = "F (foxes)") |>
  gf_refine(coord_fixed()) |>
  gf_theme(theme_minimal())
```

**Questions**

1. At the point in state space where there are *many* rabbits and *few* foxes (e.g. top-left area of the plot, high $R$, low $F$), in which direction does the flow mainly point?

```{mcq}
#| label: leh-highR-lowF
1. Mainly to the right (R increasing, F not changing much).
2. Mainly upward (F increasing, R not changing much).
3. Roughly northeast (both R and F increasing). [correct]
4. Mainly downward (F decreasing).
5. Roughly southwest (both R and F decreasing).
```

2. At the point where there are *few* rabbits and *many* foxes (e.g. bottom-right area: low $R$, high $F$), in which direction does the flow mainly point?

```{mcq}
#| label: leh-lowR-highF
1. Mainly to the right (R increasing).
2. Roughly southwest (both R and F decreasing). [correct]
3. Roughly northeast (both R and F increasing).
4. Mainly upward (F increasing).
5. Mainly to the left (R decreasing only).
```

3. What shape do the two trajectories (green and brown) in the figure have?

```{mcq}
#| label: leh-shape
1. They head straight toward the center and stop.
2. They spiral inward toward a single point.
3. They form closed loops (cycles) that repeat. [correct]
4. They head outward without bound.
5. They are straight lines.
```

4. How does this predator–prey system *differ* from the SIR epidemic model in terms of long-term behavior?

```{mcq}
#| label: leh-vs-SIR
1. SIR has no state space; predator-prey does.
2. In SIR the state eventually moves in one general direction (e.g. toward fewer infectives); in this predator-prey model the state can cycle and not settle at a single point. [correct]
3. SIR has two state variables; predator-prey has only one.
4. Predator-prey has no equilibrium; SIR always has one.
```

5. Suppose we start at a state with many rabbits and few foxes. Over the next stretch of time, what typically happens to the fox population $F$?

```{mcq}
#| label: leh-fox-increase
1. $F$ only decreases, because there are too many rabbits.
2. $F$ increases (more food for foxes), then later can decrease as rabbits are eaten down. [correct]
3. $F$ stays constant.
4. $F$ decreases to zero.
```

6. In the middle of the state space (e.g. near $R \approx 1.5$, $F \approx 2$), the arrows are relatively short. What does that indicate?

```{mcq}
#| label: leh-short-arrows
1. Time runs more slowly there.
2. The rate of change of the state is small there (state is changing slowly); this is near an equilibrium. [correct]
3. Rabbits and foxes are equal in number.
4. The trajectory cannot pass through that region.
```

7. (Optional) In one or two sentences, explain in terms of “rabbits” and “foxes” why the flow points roughly southwest when there are few rabbits and many foxes.

`r devoirs::devoirs_text("leh-explain", rows = 3)`

:::
<!-- end of exr-lynx-eat-hare -->
