---
id: "spider-put-plant"
created: "Fri Jan 16 15:16:48 2026"
attribution: TBA
---

```{r include=FALSE, eval=!"devoirs" %in% (.packages())}
# For stand-alone compilation testing
library(devoirs)
library(mosaicCalc)
```

::: {#exr-spider-put-plant}
<!-- This is a subset of maple-shut-pantry in QR-A -->

Consider the data on the body mass (kg) and brain mass (gm) for 62 species of mammals shown in @fig-mammal-mass. (You can hover the cursor over a point to see the species name.)

```{r}
#| echo: false
#| label: fig-mammal-mass
#| warning: false
library(ggformula)
library(ggiraph)
mammals <- read.csv("www/mammals.csv")
P <- mammals |> 
  ggformula:::gf_point_interactive(brain_weight ~ body_weight,
                       tooltip = paste("Species: ", mammals$species)) |>
  gf_refine(
    scale_y_log10(),
    scale_x_log10(),
    coord_fixed()
  ) |>
  gf_lm(interval = "prediction", level = 0.67) |>
  gf_labs(y = "Brain mass (gm)", x = "Body mass (kg)") |>
  gf_theme(theme_minimal())
P |> gf_girafe()
```

1. The lessor short-tailed shrew has both the smallest body mass and brain mass. What is the brain mass of the short-tailed shrew? (Pick the answer that's closest.)

```{mcq}
#| label: msp-1-4lse
#| inline: true
1. 0.03 gm
2. 0.14 gm [correct]
3. 0.52 gm
```

2. The water opposum, arctic fox, nine-banded armadillo, and rock hyrax (in the middle of the cloud of points) all have about the same body mass. What is that body mass? (Pick the answer that's closest.) 

```{mcq}
#| label: msp-2-sddo
#| inline: true
1. 1-2 kg
2. 3-4 kg [correct]
3. 6-7 kg
4. 8-9 kg
```

3. Which of these animals has a body mass closest to 10 kg?

```{mcq}
#| label: msp-3-fhsl
#| inline: true
1. raccoon
1. baboon [correct]
2. roe deer
4. vervet
```

4. Of the mammals in the range of body masses from 10 to 100 kg, which has the smallest brain size?

```{mcq}
#| label: msp-4-le4
#| inline: true
1. water opposum [hint: It's smaller than 10 kg, so not the answer to the question.]
2. giant armadillo
3. goat
4. kangaroo [correct]
```

:::
 <!-- end of exr-spider-put-plant -->
