---
number-offset: 5
crossref:
  exr-title: "Drill 6."
---

# Drill Questions: Storybook Functions {-}

```{r include=FALSE}
source("_start-up.R")
library(mosaicCalc)
```

{{< include _devoirs.qmd >}}

`r devoirs::push_answer_style("block")`

:::: {.callout-note collapse=true}
## Identifying Storybook Functions

`r cc <- drill::new_drill_counter()`
`r id <- "d06"`

::: {.panel-tabset}

## `r cc()`

Which of the following is **not** the name of a storybook function?

```{mcq}
#| label: d06-not-storybook
#| inline: true
1. hill()
2. ramp() [correct]
3. steady()
4. doublings()
5. recip()
```

## `r cc()`

Match each storybook function to its characteristic behavior. The function `osc()` ...

```{mcq}
#| label: d06-osc-behavior
1. goes up and then back down to a baseline
2. repeats the same pattern over and over [correct]
3. starts low and levels off at a high value
4. increases by the same factor for each unit increase in input
```

## `r cc()`

The function `hillside()` ...

```{mcq}
#| label: d06-hillside-behavior
1. goes up and then back down to a baseline
2. repeats the same pattern over and over
3. starts at a low value and gradually levels off at a higher value [correct]
4. decreases quickly at first, then more slowly
```

## `r cc()`

The function `hill()` ...

```{mcq}
#| label: d06-hill-behavior
1. goes up from a baseline and then back down to the baseline [correct]
2. repeats the same pattern over and over
3. starts at a low value and gradually levels off at a higher value
4. increases by the same factor for each unit increase in input
```

## `r cc()`

The function `double()` ...

```{mcq}
#| label: d06-double-behavior
1. goes up from a baseline and then back down to the baseline
2. the output doubles for every unit increase in input [correct]
3. tells you how many doublings correspond to a given output
4. decreases quickly at first, then more slowly
```

## `r cc()`

The function `doublings()` is closely related to `double()`. What is the relationship?

```{mcq}
#| label: d06-doublings-relationship
1. They are the same function with different names
2. doublings() tells you the input needed to reach a given output from double() [correct]
3. doublings() is double() multiplied by 2
4. doublings() is the rate of change of double()
```

## `r cc()`

Which storybook function would best describe the outdoor temperature over the course of a single day (rising in the morning, falling in the evening, repeating daily)?

```{mcq}
#| label: d06-temp-daily
#| inline: true
1. hill()
2. hillside()
3. osc() [correct]
4. double()
```

## `r cc()`

Which storybook function would best describe the adoption of a new technology (few users at first, rapid growth, then leveling off as the market saturates)?

```{mcq}
#| label: d06-adoption
#| inline: true
1. hill()
2. hillside() [correct]
3. osc()
4. double()
```

:::
::::

:::: {.callout-note collapse=true}
## Function Properties

`r cc <- drill::new_drill_counter(style = "ALPHA")`

::: {.panel-tabset}

## `r cc()`

Which storybook functions have a positive rate of change for all positive inputs?

`r devoirs::devoirsTF(FALSE, leed="recip()", id="d06-pos-recip")`

`r devoirs::devoirsTF(TRUE, leed="double()", id="d06-pos-double")`

`r devoirs::devoirsTF(FALSE, leed="hill()", id="d06-pos-hill")`

`r devoirs::devoirsTF(TRUE, leed="hillside()", id="d06-pos-hillside")`

`r devoirs::devoirsTF(TRUE, leed="steady()", id="d06-pos-steady")`

## `r cc()`

For the `osc()` function, what is the **period** (the input interval over which the function completes one full cycle and returns to its starting pattern)?

```{mcq}
#| label: d06-osc-period
#| inline: true
1. 0.5
2. 1 [correct]
3. 2
4. 3.14159
```

## `r cc()`

At what input value does `hillside()` produce an output of 0.5 (halfway between its minimum and maximum)?

```{mcq}
#| label: d06-hillside-half
#| inline: true
1. -1
2. 0 [correct]
3. 0.5
4. 1
```

## `r cc()`

For the `double()` function:

a. What is the output when the input is 0?

`r devoirs::devoirs_set(choices = c(0, 0.5, 1, 2, 4), correct = 1, id = "d06-double-zero")`

b. What is the output when the input is 3?

`r devoirs::devoirs_set(choices = c(3, 6, 8, 9), correct = 8, id = "d06-double-three")`

c. What is the output when the input is -1?

`r devoirs::devoirs_set(choices = c(-2, -1, 0.5, 1), correct = 0.5, id = "d06-double-neg")`

## `r cc()`

For the `doublings()` function:

a. What is the output when the input is 8?

`r devoirs::devoirs_set(choices = c(2, 3, 4, 8, 16), correct = 3, id = "d06-doublings-8")`

b. What is the output when the input is 1?

`r devoirs::devoirs_set(choices = c(-1, 0, 1, 2), correct = 0, id = "d06-doublings-1")`

c. What input to `doublings()` gives an output of -2?

`r devoirs::devoirs_set(choices = c(-4, -2, 0.25, 0.5), correct = 0.25, id = "d06-doublings-neg2")`

## `r cc()`

The `flat()` function always returns the same output regardless of input. What is the rate of change of `flat()` with respect to its input?

```{mcq}
#| label: d06-flat-rate
#| inline: true
1. -1
2. 0 [correct]
3. 1
4. It depends on the input
```

:::
::::

:::: {.callout-note collapse=true}
## Input and Output Transformations

`r cc <- drill::new_drill_counter(style = "roman")`

::: {.panel-tabset}

## `r cc()`

The storybook functions take pure numbers (dimensionless) as input and produce pure numbers as output. To use them with real-world quantities that have units, we use **input transformations**. 

An input transformation has the form $a(q - q_0)$ where $q$ is the real-world input quantity. What must be true about the dimension of $q_0$?

```{mcq}
#| label: d06-input-trans-dim
1. It must be dimensionless
2. It must have the same dimension as $q$ [correct]
3. It must have the reciprocal dimension of $q$
4. It can have any dimension
```

## `r cc()`

In the input transformation $a(q - q_0)$, what must be true about the dimension of parameter $a$?

```{mcq}
#| label: d06-a-dim
1. It must be dimensionless
2. It must have the same dimension as $q$
3. It must have dimension $1/[q]$ (the reciprocal of $q$'s dimension) [correct]
4. It can have any dimension
```

## `r cc()`

Suppose you want to model oil production using `hillside()`, and the data shows production reached half its maximum value around the year 1970. In the input transformation $a(\text{year} - q_0)$, what value should $q_0$ have?

```{mcq}
#| label: d06-q0-oil
#| inline: true
1. 0
2. 0.5
3. 1970 [correct]
4. 2023
```

## `r cc()`

An **output transformation** has the form $A \cdot f(\text{input}) + B$ where $f()$ is a storybook function. If you want the transformed function to have a maximum output of 100 and a minimum of 0, and the storybook function naturally ranges from 0 to 1, what should $A$ be?

```{mcq}
#| label: d06-output-A
#| inline: true
1. 1
2. 50
3. 100 [correct]
4. 200
```

## `r cc()`

Continuing from the previous question, what should $B$ be?

```{mcq}
#| label: d06-output-B
#| inline: true
1. -50
2. 0 [correct]
3. 50
4. 100
```

## `r cc()`

If a quantity is measured in miles per gallon and you need to create a dimensionless input for a storybook function using the transformation $a(q - q_0)$, what are the units of parameter $a$?

```{mcq}
#| label: d06-mpg-units
1. miles per gallon
2. gallons per mile [correct]
3. dimensionless
4. miles
```

:::
::::

:::: {.callout-note collapse=true}
## Recognizing Functions from Graphs

`r cc <- drill::new_drill_counter(style = "ROMAN")`

::: {.panel-tabset}

## `r cc()`

```{r}
#| echo: false
slice_plot(pnorm(x) ~ x, domain(x = -3:3)) |>
  gf_labs(y = "output", x = "input") |>
  gf_theme(theme_minimal())
```

Which storybook function does this graph show?

```{mcq}
#| label: d06-graph-hillside
#| inline: true
1. hill()
2. hillside() [correct]
3. osc()
4. double()
```

## `r cc()`

```{r}
#| echo: false
slice_plot(dnorm(x) ~ x, domain(x = -3:3)) |>
  gf_labs(y = "output", x = "input") |>
  gf_theme(theme_minimal())
```

Which storybook function does this graph show?

```{mcq}
#| label: d06-graph-hill
#| inline: true
1. hill() [correct]
2. hillside()
3. osc()
4. recip()
```

## `r cc()`

```{r}
#| echo: false
slice_plot(2^x ~ x, domain(x = -2:3)) |>
  gf_labs(y = "output", x = "input") |>
  gf_theme(theme_minimal())
```

Which storybook function does this graph show?

```{mcq}
#| label: d06-graph-double
#| inline: true
1. doublings()
2. double() [correct]
3. steady()
4. hillside()
```

## `r cc()`

```{r}
#| echo: false
slice_plot(sin(2*pi*x) ~ x, domain(x = 0:3), npts = 500) |>
  gf_labs(y = "output", x = "input") |>
  gf_theme(theme_minimal())
```

Which storybook function does this graph show?

```{mcq}
#| label: d06-graph-osc
#| inline: true
1. hill()
2. hillside()
3. osc() [correct]
4. steady()
```

## `r cc()`

```{r}
#| echo: false
slice_plot(1/x ~ x, domain(x = 0.2:3)) |>
  gf_labs(y = "output", x = "input") |>
  gf_theme(theme_minimal())
```

Which storybook function does this graph show?

```{mcq}
#| label: d06-graph-recip
#| inline: true
1. doublings()
2. double()
3. recip() [correct]
4. hillside()
```

:::
::::
